---
title: "Figure 1"
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load the libraries
library(tidyverse)
library(readr)
library(here)
library(psych)
library(english)
library(stringr)
library(tableone)
library(kableExtra)
library(scales)
library(ggridges)
library(plotly)
library(patchwork)
library(haven)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
```

```{r data, echo=FALSE, message=FALSE, warning=FALSE}
# Read in the clean csv file
leger <- read_csv("data/leger_for_analyses.csv")

# subset for chronic people only
leger_chronic <- leger %>%
  filter(chronic_dummy == 1)
```

**Figure 1: Proportion of individuals endorsing various vaccine motivators.** 
Panel A and B show the most commonly reported COVID-19 vaccine motivations among Canadians with physical chronic health conditions who were at least partially vaccinated. The question asked, “To what extent did each of the following influence your decision to get the COVID-19 vaccine?” * indicates that in the latest survey wave the wording of the question changed to ask specifically about what would motivate them to get the second dose. Likewise, panel C shows the same variables, but among unvaccinated Canadians with physical chronic health conditions. The question asked, “To what extent would the following influence your decision to get the COVID-19 vaccine?” The responses were rated on a 4-point Likert scale from To a great extend (depicted), Somewhat, Very little, Not at all.
Abbreviations: Protect others = Knowing that getting vaccinated will help protect others around me; Doing my part = Wanting to do my part to achieve "herd immunity", Info about effectiveness = Having information that the vaccine is effective (i.e., will prevent COVID-19 transmission) , Info of no side effects = Having information that the vaccine is unlikely to have any major side effects, Convenience of getting vaccine = , Reduce my worries = Believing that getting vaccinated would reduce my worries and anxiety, Trust in vaccine company = Trusting the company who developed the vaccine, Attending events = Learning that being vaccinated would allow me to attend public events (e.g., concerts, sporting events) or travel, Personal high risk = Believing that I am high risk of getting COVID-19, Others getting vaccine = Seeing more people getting the vaccine, Other’s positive att. = Hearing that other people have positive attitudes towards the vaccine, Doctor recommendation = Getting a recommendation from my doctor to get vaccinated, Employer recommendation = Getting a recommendation from my employer to get vaccinated, Only need 1 dose = Only needing one dose of the vaccine to be protected.


```{r echo=FALSE, warning=FALSE, fig.width=12, fig.height=18}

vaxed_fully <- leger_chronic %>%
  filter(vaccine == "Yes (fully)") %>%
  mutate_at(vars(c(inf1:inf16)), 
            funs(case_when(. == 1 ~ 1,
                           . == 2 | . == 3 | . == 4 ~ 0))) %>%
  pivot_longer(names_to = "influence", values_to = "value", cols = c(inf1:inf16)) %>%
  group_by(influence, value) %>%
  count(influence) %>%
  drop_na() %>%
  ungroup(influence, value) %>%
  filter(value == 1) %>%
  select(-value) %>%
  mutate(prop = n/1343*100) %>%
  mutate(prop = round(prop, 1)) %>%
  mutate(influence = as.factor(influence),
         influence = fct_reorder(influence, prop)) %>%
  ggplot(aes(x = prop, y = influence, fill = influence)) + 
  geom_col() + 
  geom_text(aes(label = paste0(prop,"%"), x = prop-3)) +
  labs(title = "Motivations among vaccinated individuals (N = 1,343)", 
       x = "Proportion of people responding 'To a great extent'",
       y = " ") + 
  xlim(0, 100) +
  theme_classic(base_size = 16) +
  theme(legend.position = "none",
        strip.text = element_text(face = 'bold'),
        panel.grid.major.x = element_line()) +
  scale_y_discrete(breaks = c("inf1", "inf2", "inf3", "inf4", "inf5", "inf6", "inf7", "inf10",
                              "inf11", "inf12", "inf13", "inf14", "inf15", "inf16", "inf17", "inf18"),
                   labels = c("Trust in vaccine company",
                              "Info of no side effects",
                              "Info about effectiveness",
                              "Personal high risk",
                              "Other's positive att.",
                              "Others getting vaccine",
                              "Doctor recommendation",
                              "Convenience of getting vaccine",
                              "Protect others",
                              "Attending events",
                              "Doing my part",
                              "Employer recommendation",
                              "Only need 1 dose",
                              "Reduce my worries",
                              "Effective for variants",
                              "Vaccine of choice")) +
  scale_fill_viridis_d(alpha = 0.9,
                       begin = 0.3, end = 1)


unvaxed_3 <- leger_chronic %>%
  filter(vaccine == "No") %>%
  mutate_at(vars(c(inf1:inf16)), 
            funs(case_when(. == 1 ~ 1,
                           . == 2 | . == 3 | . == 4 ~ 0))) %>%
  pivot_longer(names_to = "influence", values_to = "value", cols = c(inf1:inf16)) %>%
  group_by(influence, value) %>%
  count(influence) %>%
  drop_na() %>%
  ungroup(influence, value) %>%
  filter(value == 1) %>%
  select(-value)%>%
  mutate(prop = n/251*100) %>%
  mutate(prop = round(prop, 1)) %>%
  mutate(influence = as.factor(influence),
         influence = fct_reorder(influence, prop)) %>%
  ggplot(aes(x = prop, y = influence, fill = influence)) + 
  geom_col() + 
  geom_text(aes(label = paste0(prop,"%"), x = prop-3)) +
  labs(title = "Motivations among unvaccinated individuals (N = 251)", 
       x = "Proportion of people responding 'To a great extent'",
       y = " ") + 
  xlim(0, 100) +
  theme_classic(base_size = 16) +
  theme(legend.position = "none",
        strip.text = element_text(face = 'bold'),
        panel.grid.major.x = element_line()) +
  scale_y_discrete(breaks = c("inf1", "inf2", "inf3", "inf4", "inf5", "inf6", "inf7", "inf10",
                              "inf11", "inf12", "inf13", "inf14", "inf15", "inf16", "inf17", "inf18"),
                   labels = c("Trust in vaccine company",
                              "Info of no side effects",
                              "Info about effectiveness",
                              "Personal high risk",
                              "Other's positive att.",
                              "Others getting vaccine",
                              "Doctor recommendation",
                              "Convenience of getting vaccine",
                              "Protect others",
                              "Attending events",
                              "Doing my part",
                              "Employer recommendation",
                              "Only need 1 dose",
                              "Reduce my worries",
                              "Effective for variants",
                              "Vaccine of choice")) +
  scale_fill_viridis_d(alpha = 0.9,
                       begin = 0.3, end = 1)

vaxed_partially <- leger_chronic %>%
  filter(vaccine == "Yes (partially)") %>%
  mutate_at(vars(c(inf1:inf16)), 
            funs(case_when(. == 1 ~ 1,
                           . == 2 | . == 3 | . == 4 ~ 0))) %>%
  pivot_longer(names_to = "influence", values_to = "value", cols = c(inf1:inf16)) %>%
  group_by(influence, value) %>%
  count(influence) %>%
  drop_na() %>%
  ungroup(influence, value) %>%
  filter(value == 1) %>%
  select(-value)%>%
  mutate(prop = n/951*100) %>%
  mutate(prop = round(prop, 1)) %>%
  mutate(influence = as.factor(influence),
         influence = fct_reorder(influence, prop)) %>%
  ggplot(aes(x = prop, y = influence, fill = influence)) + 
  geom_col() + 
  geom_text(aes(label = paste0(prop,"%"), x = prop-3)) +
  labs(title = "Motivations among partially vaccinated individuals (N = 951)*", 
       x = "Proportion of people responding 'To a great extent'",
       y = " ") + 
  xlim(0, 100) +
  theme_classic(base_size = 16) +
  theme(legend.position = "none",
        strip.text = element_text(face = 'bold'),
        panel.grid.major.x = element_line()) +
  scale_y_discrete(breaks = c("inf1", "inf2", "inf3", "inf4", "inf5", "inf6", "inf7", "inf10",
                              "inf11", "inf12", "inf13", "inf14", "inf15", "inf16", "inf17", "inf18"),
                   labels = c("Trust in vaccine company",
                              "Info of no side effects",
                              "Info about effectiveness",
                              "Personal high risk",
                              "Other's positive att.",
                              "Others getting vaccine",
                              "Doctor recommendation",
                              "Convenience of getting vaccine",
                              "Protect others",
                              "Attending events",
                              "Doing my part",
                              "Employer recommendation",
                              "Only need 1 dose",
                              "Reduce my worries",
                              "Effective for variants",
                              "Vaccine of choice")) +
  scale_fill_viridis_d(alpha = 0.9,
                       begin = 0.3, end = 1)

(vaxed_fully/vaxed_partially/unvaxed_3) + plot_annotation(tag_levels = 'A')
```

