---
title: "Canada: COVID-19 vaccine and booster attitudes and motivations in adults with physical chronic health conditions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load the libraries
library(tidyverse)
library(readr)
library(here)
library(psych)
library(english)
library(stringr)
library(tableone)
library(kableExtra)
library(scales)
library(ggridges)
library(plotly)
library(patchwork)
library(haven)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
```

```{r data, echo=FALSE, message=FALSE, warning=FALSE}
# Read in the clean csv file
leger <- read_csv("data/leger_for_analyses.csv")

# subset for chronic people only
leger_chronic <- leger %>%
  filter(chronic_dummy == 1)
```
## INTRODUCTION

Covid-19 vaccines have become wide-spread in the Western world beginning of 2021. To date, Canada's vaccine uptake is among leading in the world, with over 75% of eligible population vaccinated.

#### Aims:
Using Canadian representative samples, we aimed 

1) to explore attitudes about the COVID-19 vaccine booster doses among the Canadian adult population with and without physical chronic health conditions, and 

2) to understand what are the motivations for vaccine uptake and factors associated with COVID-19 vaccine resistance among individuals with physical chronic health conditions. 

## METHOD

The present study is part of a larger International COVID‐19 Awareness and Responses Evaluation (iCARE) study, which is a global multi‐phase research survey conducted through the Montreal Behavioural Medicine Centre. Find more about the study and materials at [www.iCAREstudy.com](www.iCAREstudy.com). 

## RESULTS
### Sample

NB - this automatic demographic table does not contain weights therefore the differences between the groups (t-test and chi2) will be recalculated.

More complete table is available in the separate file on the same repository.
```{r echo=FALSE, warning=FALSE}
## Create a table
table_data <- leger %>%
  select(age_yrs, sex, edu, hoinc, ethnicity, province_full, chronic_descriptive, chronic_dummy, chronic_multi, any_m_health)

# First create a variable list you want in the table
var_list <- c("age_yrs", "sex", "edu", "hoinc", "province_full", "chronic_descriptive", "chronic_multi")

# Determine which variables are categorical
cat_vars <- c("sex", "edu", "hoinc", "province_full", "chronic_descriptive", "chronic_multi")

#All cases table
table_1 <- CreateTableOne(vars = var_list,
                          factorVars = cat_vars,
                          data = table_data, 
                          strata = c("chronic_dummy"))

table_1
```

### Vaccine uptake and booster doses
When broken down by the health status, we found that only 9.8% of people with physical chronic health conditions reported being non-vaccinated, compared to 14.3% of healthy individuals.

```{r echo=FALSE, warning=FALSE}
leger %>%
  select(chronic_dummy, vaccine) %>%
  group_by(chronic_dummy, vaccine) %>%
  drop_na() %>%
  count() %>%
  group_by(chronic_dummy) %>%
  mutate(sum_group_status = sum(n),
         prop = n / sum_group_status * 100,
         prop = round(prop, 2)) %>%
  select(-sum_group_status) %>%
  kable(col.names = c("Chronic illness", "Vaccine status", "N", "Proportion")) %>%
  kable_styling(full_width = FALSE)
```

Canadian with chronic health conditions demonstrate positive attitudes toward COVID-19 vaccine boosters, only 5% reported being somewhat or extremely unlikely to take the booster dose, compared to 10% of health individuals.
```{r echo=FALSE, warning=FALSE}
leger %>%
  select(chronic_dummy, booster_dummy) %>%
  group_by(chronic_dummy, booster_dummy) %>%
  drop_na() %>%
  count() %>%
  group_by(chronic_dummy) %>%
  mutate(sum_group_status = sum(n),
         prop = n / sum_group_status * 100,
         prop = round(prop, 2)) %>%
  select(-sum_group_status) %>%
  kable(col.names = c("Chronic illness", "Vaccine status", "N", "Proportion")) %>%
  kable_styling(full_width = FALSE)
```

### Vaccine motivations among Canadian with physical chronic health conditions

```{r echo=FALSE, warning=FALSE, fig.width=12, fig.height=14}
vaxed <- leger_chronic %>%
  filter(vaccine_dummy == 1) %>%
  mutate_at(vars(c(inf1:inf16)), 
            funs(case_when(. == 1 ~ 1,
                           . == 2 | . == 3 | . == 4 ~ 0))) %>%
  pivot_longer(names_to = "influence", values_to = "value", cols = c(inf1:inf16)) %>%
  group_by(influence, value) %>%
  count(influence) %>%
  drop_na() %>%
  ungroup(influence, value) %>%
  filter(value == 1) %>%
  select(-value) %>%
  mutate(influence = as.factor(influence),
         influence = fct_reorder(influence, n)) %>%
  ggplot(aes(x = n, y = influence, fill = influence)) + 
  geom_col() + 
  geom_text(aes(label = paste0(round((n/2294*100),1),"%"), x = n - 70)) +
  labs(title = "Motivations among vaccinated individuals (N = 2,294)", 
       x = "Number of people responding 'To a great extent'",
       y = " ") + 
  theme_classic(base_size = 16) +
  theme(legend.position = "none",
        strip.text = element_text(face = 'bold'),
        panel.grid.major.x = element_line()) +
  scale_y_discrete(breaks = c("inf1", "inf2", "inf3", "inf4", "inf5", "inf6", "inf7", "inf10",
                              "inf11", "inf12", "inf13", "inf14", "inf15", "inf16", "inf17", "inf18"),
                   labels = c("Trust in vaccine company",
                              "Info of no side effects",
                              "Info about effectiveness",
                              "Personal high risk",
                              "Other's positive att.",
                              "Others getting vaccine",
                              "Doctor recommendation",
                              "Convenience of getting vaccine",
                              "Protect others",
                              "Attending events",
                              "Doing my part",
                              "Employer recommendation",
                              "Only need 1 dose",
                              "Reduce my worries",
                              "Effective for variants",
                              "Vaccine of choice")) +
  scale_fill_viridis_d(alpha = 0.9,
                       begin = 0.3, end = 1)


unvaxed <- leger_chronic %>%
  filter(vaccine_dummy == 0) %>%
  mutate_at(vars(c(inf1:inf16)), 
            funs(case_when(. == 1 ~ 1,
                           . == 2 | . == 3 | . == 4 ~ 0))) %>%
  pivot_longer(names_to = "influence", values_to = "value", cols = c(inf1:inf16)) %>%
  group_by(influence, value) %>%
  count(influence) %>%
  drop_na() %>%
  ungroup(influence, value) %>%
  filter(value == 1) %>%
  select(-value) %>%
  mutate(influence = as.factor(influence),
         influence = fct_reorder(influence, n)) %>%
  ggplot(aes(x = n, y = influence, fill = influence)) + 
  geom_col() + 
  geom_text(aes(label = paste0(round((n/251*100),1),"%"), x = n - 5)) +
  labs(title = "Motivations among unvaccinated individuals (N = 251)", 
       x = "Number of people responding 'To a great extent'",
       y = " ") + 
  theme_classic(base_size = 16) +
  theme(legend.position = "none",
        strip.text = element_text(face = 'bold'),
        panel.grid.major.x = element_line()) +
  scale_y_discrete(breaks = c("inf1", "inf2", "inf3", "inf4", "inf5", "inf6", "inf7", "inf10",
                              "inf11", "inf12", "inf13", "inf14", "inf15", "inf16", "inf17", "inf18"),
                   labels = c("Trust in vaccine company",
                              "Info of no side effects",
                              "Info about effectiveness",
                              "Personal high risk",
                              "Other's positive att.",
                              "Others getting vaccine",
                              "Doctor recommendation",
                              "Convenience of getting vaccine",
                              "Protect others",
                              "Attending events",
                              "Doing my part",
                              "Employer recommendation",
                              "Only need 1 dose",
                              "Reduce my worries",
                              "Effective for variants",
                              "Vaccine of choice")) +
  scale_fill_viridis_d(alpha = 0.9,
                       begin = 0.3, end = 1)

(vaxed/unvaxed) + plot_annotation(tag_levels = 'A')
```

### Logistic regression output

*NOTE* - Education is treated as continuous variable with the following levels: "never been" ~ 0, "primary" ~ 1,"secondary" ~ 2, "college" ~ 3, "graduate" ~ 4.

Province is broken down into 3 geographical regions: 

* *East* = "New Brunswick", "Nova Scotia", "Prince Edward Island", "Newfoundland"
* *Central* = "Ontario", "Quebec"
* *West* = "British Columbia", "Alberta", "Saskatchewan", "Manitoba"

Distress rev = compound score of COVID-19 related feelings of anxiety, depression, and loneliness. Scale rangess from 0 - 9, with higher scores indicating higher levels of pandemic-related distress.

```{r echo=FALSE, warning=FALSE}
set.seed(1)

logit_model_data <- leger_chronic %>%
  select(vaccine_dummy, age_yrs, sex, hoinc, edu, province, wave, any_m_health, chronic_multi, distress_rev, pond) %>%
  filter(sex %in% c("Male", "Female")) %>%
  mutate(education = as.numeric(case_when(edu == "Never been" ~ 0, 
                                          edu == "Primary" ~ 1,
                                          edu == "Secondary" ~ 2,
                                          edu == "College" ~ 3,
                                          edu == "Graduate" ~ 4)))


# In case income should be treated as numeric
# income = as.numeric(case_when(hoinc == "bottom 3rd" ~ 0,
#                                        hoinc == "middle 3rd" ~ 1,
#                                        hoinc == "top 3rd" ~ 2)),

logit_c1 <- glm(vaccine_dummy ~ age_yrs + sex + province + education + hoinc + any_m_health + chronic_multi + distress_rev,
                family = "binomial",
                data = logit_model_data,
                weights = pond)

tab_model(logit_c1)

```

## DISCUSSION

These insights provide important information on how healthcare professionals, policymakers as well as all the relevant stakeholders may need to structure their vaccine messaging in order to encourage the vaccine uptake among the most vulnerable and at risk populations.

